### 2011-09-10 paper.tex
deleted old text relating to SLIGs
Better notation/formalism

\Tree[ .$\epsilon$ [ .$\startsymbol$ [ .$\epsilon$ ] ] ]

@@

Each ``intermediate tree'' is, formally, an ordered tree whose nodes are labeled from
$(\nodelabels \cup \terminals^\ast)$.
The transformation rules can take the various forms shown in \tabref{RuleTypes}.

@@

To obtain the final output sequence, we read off the terminal sequences at the leaves of the tree,
from just left of the root, moving anticlockwise around the tree, to just right of the root.

@@

The {\em weight} of a derivation is the product of the weights of all rules used in the derivation.
The weight of a given output sequence $\outseq$ is the sum of the weights of all possible derivations of $\outseq$.
Let $\seqweight{\cal T}{\outseq}$ denote the weight of output sequence $\outseq$,
starting from initial tree ${\cal T}$.

The final column of \tabref{RuleTypes} shows a shorthand representation of each rule in Newick format,
probably the most widely-understood bioinformatics format for representing tree structures.
The initial tree, in this representation, is $((\epsilon)S)\epsilon$.
In the table, we have omitted the placeholder $\epsilon$'s at leaf nodes,
so that (for example) the rule
\[
((\beta)A)\alpha\ \to\ ((C,\beta)B)\alpha
\]
should strictly be read as
\[
((\beta)A)\alpha\ \to\ (((\epsilon)C,\beta)B)\alpha
\]

Let $\rulesubset{n}{A} \subseteq \rules$ denote the subset of rules of type $n$ (according to \tabref{RuleTypes})
where the left-hand side is $((\beta)A)\alpha$.

Note the special case of a type-6 rule where all the terminal strings are empty,
$((\beta)A)\alpha \to ((\beta)B)\alpha$.
Such a rule is referred to as a {\em transition} and may be written more compactly as $A \to B$.

\begin{table}
\tiny
\begin{tabular}{m{.2in}m{1in}m{.2in}m{1in}r}
\centerline{Type} & \centerline{From} & & \centerline{To} & Newick representation
\\ \hline
(1) &
\Tree[ .$\alpha$ [ .$A$ [ .$\beta$ ] ] ]
 & $\to$ & 
\Tree[ .$\alpha$ [ .$B$ [ [ [ .$C$ $\epsilon$ ] ] [ [ .$\beta$ ] ] ] ] ]
& $((\beta)A)\alpha \to ((C,\beta)B)\alpha$
\\ \hline
(2) &
\Tree[ .$\alpha$ [ .$A$ [ .$\beta$ ] ] ]
 & $\to$ & 
\Tree[ .$\alpha$ [ .$B$ [ [ [ .$\beta$ ] ] [ [ .$C$ $\epsilon$ ] ] ] ] ]
& $((\beta)A)\alpha \to ((\beta,C)B)\alpha$
\\ \hline
(3) &
\Tree[ .$\alpha$ [ .$A$ [ .$\beta$ ] ] ]
 & $\to$ & 
\Tree[ .$\alpha$ [ [ [ .$C$ $\epsilon$ ] ] [ [ .$B$ $\beta$ ] ] ] ]
& $((\beta)A)\alpha \to (C,((\beta)B))\alpha$
\\ \hline
(4) &
\Tree[ .$\alpha$ [ .$A$ [ .$\beta$ ] ] ]
 & $\to$ & 
\Tree[ .$\alpha$ [ [ [ .$B$ $\beta$ ] ] [ [ .$C$ $\epsilon$ ] ] ] ]
& $((\beta)A)\alpha \to (((\beta)B),C)\alpha$
\\ \hline
(5) &
\Tree[ .$\alpha$ [ .$A$ [ .$\beta$ ] ] ]
 & $\to$ & 
\Tree[ .$\alpha$ [ .$B$ [ .$C$ $\beta$ ] ] ]
& $((\beta)A)\alpha \to (((\beta)C)B)\alpha$
\\ \hline
(6) &
\Tree[ .$\alpha$ [ .$A$ [ .$\beta$ ] ] ]
 & $\to$ & 
\Tree[ .$\alpha$ [ $u$ [ .$B$ [ $v$ [ $\beta$ ] $w$ ] ] $x$ ] ]
& $((\beta)A)\alpha \to (u,((v,\beta,w)B),x)\alpha$
\\ \hline
\end{tabular}
\normalsize
\caption{
\tablabel{RuleTypes}
Types of transformation rule (i.e. tree adjunction rule) used in this paper.
Here $\alpha,\beta$ represent any subtree;
$A \in \nodelabels$ is the source node label;
$B,C \in (\nodelabels \cup \{\epsilon\})$ are the destination node labels;
$\epsilon$ is the empty string;
and
$u,v,w,x \in \terminals^\ast$ are (possibly empty) terminal strings.
}
\end{table}

@@

Define some indicator functions to match output substrings to rules
\begin{eqnarray*}
\Delta(i,j,x) & = & \delta \left( \outsubseq{i}{j}\ =\ x \right) \\
\Delta(i,x) & = & \Delta(i,i+|x|,x)
\end{eqnarray*}

@@

Introducing the placeholder $\gamma$ as an additional terminal, define
\begin{eqnarray*}
\m(i,j,k,l,X) & = & \seqweight{((\gamma)X)\epsilon}{\outsubseq{i}{j} \gamma \outsubseq{k}{l}} \\
\mtip(i,l,X) & = & \seqweight{((\epsilon)X)\epsilon}{\outsubseq{i}{l}}
\end{eqnarray*}
where $1 \leq i \leq j \leq k \leq l \leq |\outseq|$.

@@

Define the following boundary conditions:
\begin{eqnarray*}
\m(i,i,k,k,\epsilon) & = & 1 \\
\mtip(i,i,\epsilon) & = & 1
\end{eqnarray*}

@@

The DP recursion is as follows (with $B,C,u,v,w,x$ defined as in \tabref{RuleTypes})
\begin{eqnarray*}
\lefteqn{\m(i,j,k,l,A) =} \\
& &
\sum_{\rho \in \rulesubset{1}{A}} \sum_{m=i}^j W(\rho) \m(m,j,k,l,B) \mtip(i,m,C) \\
& & +
\sum_{\rho \in \rulesubset{2}{A}} \sum_{n=k}^l W(\rho) \m(i,j,k,n,B) \mtip(n,l,C) \\
& & +
\sum_{\rho \in \rulesubset{3}{A}} \sum_{m=i}^j W(\rho) \m(i,m,k,l,B) \mtip(m,j,C) \\
& & +
\sum_{\rho \in \rulesubset{4}{A}} \sum_{n=k}^l W(\rho) \m(i,j,n,l,B) \mtip(k,n,C) \\
& & +
\sum_{\rho \in \rulesubset{5}{A}} \sum_{m=i}^j \sum_{n=k}^l W(\rho) \m(i,m,n,l,B) \m(m,j,k,n,C) \\
& & +
\sum_{\rho \in \rulesubset{6}{A}} W(\rho) \m(i+|u|,j-|v|,k+|w|,l-|x|,B)
 \Delta(i,u) \Delta(j-|v|,v) \Delta(k,w) \Delta(l-|x|,x)
\end{eqnarray*}

Note that for $\m$ to be exactly computable, we require that there are no {\em null cycles} in the grammar.
A null cycle is a series of transformations that, when applied consecutively to a given tree ${\cal T}$, yield the original tree ${\cal T}$ again,
or one that is trivially related to it (e.g. a tree that is identical to ${\cal T}$ after internal $\epsilon$-labeled nodes have been removed).
Null cycles frequently arise via consecutive transitions of the form $A \to B$ followed by $B \to A$,
but there are other possibilities too (for example, a type-1 rule $((\beta)A)\alpha \to ((C,\beta)B)\alpha$
followed by two transitions $C \to \epsilon$ and $B \to A$).

@@

@@

Note a flaw of the framework is that, since the index string representing the ``consensus'' LTR must be directly observed at least once in the output
(in fact, it corresponds to the 5'-most repeat in this grammar), 
we cannot allow a nested transposon insertion {\em within} that LTR.

@@

\noindent
\begin{tabular}{ll}
Symbol & Meaning \\
\hline
$\grammar$ & Grammar \\
$\nodelabels$ & Set of node labels \\
$\terminals$ & Set of terminals \\
$\rules$ & Set of transformation rules \\
$\rulesubset{n}{A}$ & Set of transformation rules of type $n$ (see \tabref{RuleTypes}) whose LHS is $((\beta)A)\alpha$ \\
$\alpha,\beta$ & Arbitrary ordered node-labeled trees \\
$\weight$ & Rule weight function \\
$\seqweight{\cal T}{\outseq}$ & Sum, over derivations of $\outseq$ from ${\cal T}$, of product over rule weights \\
$\epsilon$ & The empty string \\
$\terminals^\ast$ & Set of strings over $\terminals$, including the empty string \\
$\outseq$ & Output sequence \\
$|\outseq|$ & Length of $\outseq$ \\
$\outsubseq{i}{j+1}$ & Substring of $\outseq$ from $i$ to $j$ inclusive ($i$ starts at 1) \\
$\outsubseq{i}{i}$ & The empty string \\
$\Delta(i,j,x)$ & Indicates if rule string $x$ matches output string $\outsubseq{i}{j}$ \\
$\Delta(i,x)$ & Indicates if rule string $x$ matches output string $\outseq$, starting at position $i$ \\
$\m(i,j,k,l,A)$ & Inside weight for $\outsubseq{i}{j}$ (left) and $\outsubseq{k}{l}$ (right) rooted at $A$ \\
$\mtip(i,l,A)$ & Inside weight for $\outsubseq{i}{l}$ rooted at $A$ \\
\end{tabular}

>>>>

\newcommand\lineartree[3]{\Tree[ .$#1$ [ .$#2$ [ .$#3$ ] ] ]}
\newcommand\inittree[1]{\lineartree{\epsilon}{#1}{\epsilon}}

\inittree{\startsymbol}

@@

\newcommand\abtree[1]{\lineartree{\alpha}{#1}{\beta}}
\newcommand\lhs{\abtree{A}}

%\newcommand\bigrhsA{\Tree[ .$\alpha$ [ .$B$ [ [ [ .$C$ $\epsilon$ ] ] [ [ .$\beta$ ] ] ] ] ]}
%\newcommand\bigrhsB{\Tree[ .$\alpha$ [ .$B$ [ [ [ .$\beta$ ] ] [ [ .$C$ $\epsilon$ ] ] ] ] ]}
%\newcommand\bigrhsC{\Tree[ .$\alpha$ [ [ [ .$C$ $\epsilon$ ] ] [ [ .$B$ $\beta$ ] ] ] ]}
%\newcommand\bigrhsD{\Tree[ .$\alpha$ [ [ [ .$B$ $\beta$ ] ] [ [ .$C$ $\epsilon$ ] ] ] ]}
%\newcommand\bigrhsE{\Tree[ .$\alpha$ [ .$B$ [ .$C$ $\beta$ ] ] ]}
%\newcommand\bigrhsF{\Tree[ .$\alpha$ [ $u$ [ .$B$ [ $v$ [ $\beta$ ] $x$ ] ] $y$ ] ]}

\newcommand\tinyrhsA{\Tree[ .$\alpha$ [ .$C$ [ .$U$ $\epsilon$ ] [ .$\beta$ ] ] ]}
\newcommand\tinyrhsB{\Tree[ .$\alpha$ [ .$C$ [ .$\beta$ ] [ .$Y$ $\epsilon$ ] ] ]}
\newcommand\tinyrhsC{\Tree[ .$\alpha$ [ .$V$ $\epsilon$ ] [ .$C$ $\beta$ ] ]}
\newcommand\tinyrhsD{\Tree[ .$\alpha$ [ .$C$ $\beta$ ] [ .$X$ $\epsilon$ ] ]}
\newcommand\tinyrhsE{\Tree[ .$\alpha$ [ .$B$ [ .$C$ $\beta$ ] ] ]}
%\newcommand\tinyrhsF{\Tree[ .$\alpha$ $u$ [ .$C$ $v$ $\beta$ $x$ ] $y$ ]}
\newcommand\tinyrhsF{\abtree{C}}

%\newcommand\tinynewickA{((\beta)A)\alpha \to ((U,\beta)C)\alpha}
%\newcommand\tinynewickB{((\beta)A)\alpha \to ((\beta,Y)C)\alpha}
%\newcommand\tinynewickC{((\beta)A)\alpha \to (V,((\beta)C))\alpha}
%\newcommand\tinynewickD{((\beta)A)\alpha \to (((\beta)C),X)\alpha}
%\newcommand\tinynewickE{((\beta)A)\alpha \to (((\beta)C)B)\alpha}
%\newcommand\tinynewickF{((\beta)A)\alpha \to (u,((v,\beta,x)C),y)\alpha}

\newcommand\rhs[6]{\Tree[ .$\alpha$ [ .$#1$ [ .$#3$ $\epsilon$ ] [ .$#2$ [ .$#4$ $\epsilon$ ] $\beta$ [ .$#5$ $\epsilon$ ] ] [ .$#6$ $\epsilon$ ] ] ]}
\newcommand\rhsnob[5]{\Tree[ .$\alpha$ $#2$ [ .$#1$ $#3$ $\beta$ $#4$ ] $#5$ ]}
\newcommand\rhsnouvxy[2]{\Tree[ .$\alpha$ [ .$#1$ [ .$#2$ $\beta$ ] ] ]}

\newcommand\rhsGeneric{\rhs{B}{C}{U}{V}{X}{Y}}

\newcommand\rhsA{\Tree[ .$\alpha$ [ .$U$ $\epsilon$ ] [ .$c$ $v$ $\beta$ $x$ ] $y$ ]}
\newcommand\rhsB{\Tree[ .$\alpha$ $u$ [ .$c$ $v$ $\beta$ $x$ ] [ .$Y$ $\epsilon$ ] ]}
\newcommand\rhsC{\Tree[ .$\alpha$ $u$ [ .$c$ [ .$V$ $\epsilon$ ] $\beta$ $x$ ] $y$ ]}
\newcommand\rhsD{\Tree[ .$\alpha$ $u$ [ .$c$ $v$ $\beta$ [ .$X$ $\epsilon$ ] ] $y$ ]}
\newcommand\rhsE{\rhs{B}{C}{u}{v}{x}{y}}
\newcommand\rhsF{\rhsnob{C}{u}{v}{x}{y}}

\newcommand\newickA{((\beta)A)\alpha \to (U,((v,\beta,x)C),y)\alpha}
\newcommand\newickB{((\beta)A)\alpha \to (u,((v,\beta,x)C),Y)\alpha}
\newcommand\newickC{((\beta)A)\alpha \to (u,((V,\beta,x)C),y)\alpha}
\newcommand\newickD{((\beta)A)\alpha \to (u,((v,\beta,X)C),y)\alpha}
\newcommand\newickE{((\beta)A)\alpha \to ((u,((v,\beta,x)C),y)B)\alpha}
\newcommand\newickF{((\beta)A)\alpha \to (u,((v,\beta,x)C),y)\alpha}

\newcommand\longnewickA{((\beta)A)\alpha \to ((\epsilon)U,((v,\beta,x)C),y)\alpha}

Each ``intermediate tree'' is, formally, an ordered tree whose nodes are labeled from
$(\nodelabels \cup \terminals^\ast)$.
The transformation rules can take the various forms shown in \tabref{RuleTypes}.
All the rules may be recognized as special cases of the general form

\begin{tabular}{m{1in}m{.2in}m{1in}}
\lhs & $\to$ & \rhsGeneric
\end{tabular}

where $U,V,X,Y \in (\nodelabels \cup \terminals^\ast)$
and $B,C \in (\nodelabels \cup \{\epsilon\}^\ast)$.
The tree is ordered left-to-right.

@@

To obtain the final output sequence, we read off the terminal sequences (at the leaves) in a depth-first traversal of the ordered tree
(i.e. from just left of the root, moving anticlockwise around the tree, to just right of the root).
The depth-first order of leaves in the general tree shown above is $(U,V,\beta,X,Y)$.

@@

The {\em weight} of a derivation is the product of the weights of all rules used in the derivation.
The weight of a given output sequence $\outseq$ is the sum of the weights of all possible derivations of $\outseq$.
Let $\seqweight{\alpha}{\outseq}$ denote the weight of output sequence $\outseq$,
starting from initial tree $\alpha$.

The final column of \tabref{RuleTypes} shows a shorthand representation of each rule in Newick format,
probably the most widely-understood bioinformatics format for representing tree structures.
The initial tree, in this representation, is $((\epsilon)S)\epsilon$.
In the table, we have omitted the placeholder $\epsilon$'s at leaf nodes,
so that (for example) the rule
\[
\newickA
\]
should strictly be read as
\[
\longnewickA
\]

Let $\rulesubset{n}{A}$ denote the subset of rules of type $n$ (according to \tabref{RuleTypes})
where the left-hand side is $((\beta)A)\alpha$ for a given $A \in \nodelabels$.

Note the special case of a type-6 rule where all the terminal strings are empty,
$((\beta)A)\alpha \to ((\beta)C)\alpha$.
Such a rule is referred to as a {\em transition} and may be written more compactly as $A \to C$.

\begin{table}
\tiny
\begin{tabular}{m{.2in}m{.2in}m{.2in}m{1in}cm{1.2in}}
\centerline{Type} & \centerline{From} & & \centerline{To} & Newick representation & When $uvxy=\epsilon$:
\\ \hline
(1) &
\lhs & $\to$ & \rhsA & $\newickA$ & \tinyrhsA
\\ \hline
(2) &
\lhs & $\to$ & \rhsB & $\newickB$ & \tinyrhsB
\\ \hline
(3) &
\lhs & $\to$ & \rhsC & $\newickC$ & \tinyrhsC
\\ \hline
(4) &
\lhs & $\to$ & \rhsD & $\newickD$ & \tinyrhsD
\\ \hline
(5) &
\lhs & $\to$ & \rhsE & $\newickE$ & \tinyrhsE
\\ \hline
(6) &
\lhs & $\to$ & \rhsF & $\newickF$ & \tinyrhsF
\\ \hline
\end{tabular}
\normalsize
\caption{
\tablabel{RuleTypes}
Types of transformation rule (i.e. tree adjunction rule) used in this paper.
Here $\alpha,\beta$ represent any subtree;
$A \in \nodelabels$ is the source node label;
$B,C,U,V,X,Y \in (\nodelabels \cup \{\epsilon\})$ are the destination node labels;
$\epsilon$ is the empty string;
and
$u,v,x,y \in \terminals^\ast$ are (possibly empty) terminal strings.
The final column shows the destination tree when all the terminal strings $u,v,x,y$ are empty.
}
\end{table}

@@

\newcommand\match{\Delta}
\newcommand\fmatch{\stackrel{\rightarrow}{\match}}
\newcommand\rmatch{\stackrel{\leftarrow}{\match}}
\newcommand\iumatch{\match_{iu}}
\newcommand\jvmatch{\match_{jv}}
\newcommand\kxmatch{\match_{kx}}
\newcommand\lymatch{\match_{ly}}

Define some indicator functions to match output substrings to rules
\begin{eqnarray*}
\match(i,j,x) & = & \delta \left( \outsubseq{i}{j}\ =\ x \right) \\
\fmatch(i,x) & = & \match(i,i+|x|,x) \\
\rmatch(j,y) & = & \match(j-|y|,j,y) \\
\iumatch & = & \fmatch(i,u) \\
\jvmatch & = & \rmatch(j,v) \\
\kxmatch & = & \fmatch(k,x) \\
\lymatch & = & \rmatch(l,y)
\end{eqnarray*}

@@

Introducing the placeholder $\gamma$ as an additional terminal, define
\begin{eqnarray*}
\m(i,j,k,l,A) & = & \seqweight{((\gamma)A)\epsilon}{\outsubseq{i}{j} \gamma \outsubseq{k}{l}} \\
\mtip(i,l,A) & = & \seqweight{((\epsilon)A)\epsilon}{\outsubseq{i}{l}}
\end{eqnarray*}
where $1 \leq i \leq j \leq k \leq l \leq |\outseq|$.

@@

Note also the following boundary conditions:
\begin{eqnarray*}
\m(i,i,k,k,\epsilon) & = & 1 \\
\mtip(i,i,\epsilon) & = & 1
\end{eqnarray*}

@@

The DP recursion is as follows (with $B,C,U,V,X,Y,u,v,x,y$ defined as in \tabref{RuleTypes})
\begin{eqnarray*}
\lefteqn{\m(i,j,k,l,A) =} \\
& &
\sum_{\rho \in \rulesubset{1}{A}} \sum_{m=i}^j W(\rho) \m(m,j-|v|,k+|x|,l-|y|,C) \mtip(i,m,U) \jvmatch \kxmatch \lymatch \\
& & +
\sum_{\rho \in \rulesubset{2}{A}} \sum_{n=k}^l W(\rho) \m(i+|u|,j-|v|,k+|x|,n,C) \mtip(n,l,Y) \iumatch \jvmatch \kxmatch \\
& & +
\sum_{\rho \in \rulesubset{3}{A}} \sum_{m=i}^j W(\rho) \m(i+|u|,m,k+|x|,l-|y|,C) \mtip(m,j,V) \iumatch \kxmatch \lymatch \\
& & +
\sum_{\rho \in \rulesubset{4}{A}} \sum_{n=k}^l W(\rho) \m(i+|u|,j-|v|,n,l-|y|,C) \mtip(k,n,X) \iumatch \jvmatch \lymatch \\
& & +
\sum_{\rho \in \rulesubset{5}{A}} \sum_{m=i}^j \sum_{n=k}^l W(\rho) \m(i+|u|,m,n,l-|y|,B) \m(m,j-|v|,k+|x|,n,C) \iumatch \jvmatch \kxmatch \lymatch \\
& & +
\sum_{\rho \in \rulesubset{6}{A}} W(\rho) \m(i+|u|,j-|v|,k+|x|,l-|y|,C)
 \iumatch \jvmatch \kxmatch \lymatch
\end{eqnarray*}

Note that for $\m$ to be exactly computable, we require that there are no {\em null cycles} in the grammar.
A null cycle is a series of transformations that, when applied consecutively to a given tree $\alpha$, yield the original tree $\alpha$ again,
or one that is trivially related to it (e.g. a tree that is identical to $\alpha$ after $\epsilon$-labeled nodes of degree $<3$ have been removed).
Null cycles frequently arise via consecutive transitions of the form $A \to B$ followed by $B \to A$,
but there are other possibilities too (for example, a type-1 bifurcation rule $((\beta)A)\alpha \to ((U,\beta)C)\alpha$
followed by two transitions $U \to \epsilon$ and $C \to A$).

@@

The LTRs can be generated by type-6 rules (e.g. with $v=y=\epsilon$) followed either by a transition to the transposon interior,
or (optionally) a type-5 adjunction rule, which may be an easier event on which to impose a hint-requirement constraint.

@@

@@

\noindent
\begin{tabular}{ll}
Symbol & Meaning \\
\hline
$\grammar$ & Grammar \\
$\nodelabels$ & Set of node labels \\
$\terminals$ & Set of terminals \\
$\rules$ & Set of transformation rules \\
$\rulesubset{n}{A}$ & Set of transformation rules of type $n$ (see \tabref{RuleTypes}) whose LHS is $((\beta)A)\alpha$ \\
$\alpha,\beta$ & Arbitrary ordered node-labeled trees \\
$\weight$ & Rule weight function \\
$\seqweight{\alpha}{\outseq}$ & Sum, over derivations of $\outseq$ from $\alpha$, of product over derivation rule weights \\
$\epsilon$ & The empty string \\
$\terminals^\ast$ & Set of strings over $\terminals$, including the empty string \\
$\outseq$ & Output sequence \\
$|\outseq|$ & Length of $\outseq$ \\
$\outsubseq{i}{j+1}$ & Substring of $\outseq$ from $i$ to $j$ inclusive ($i$ starts at 1) \\
$\outsubseq{i}{i}$ & The empty string \\
$\match(i,j,x)$ & Indicates if rule string $x$ matches output string $\outsubseq{i}{j}$ \\
$\fmatch(i,x)$ & Indicates if rule string $x$ matches output string $\outseq$, starting at position $i$ \\
$\rmatch(j,y)$ & Indicates if rule string $y$ matches output string $\outseq$, ending before position $j$ \\
$\iumatch,\jvmatch,\kxmatch,\lymatch$ & Rule string matchers; $\match_{ns}$ matches string $s$ ending/starting at position $n$ \\
$\m(i,j,k,l,A)$ & Inside weight for $\outsubseq{i}{j}$ (left) and $\outsubseq{k}{l}$ (right) rooted at $A$ \\
$\mtip(i,l,A)$ & Inside weight for $\outsubseq{i}{l}$ rooted at $A$ \\
$B,C,U,V,X,Y$ & Labels (or terminal strings) at nodes of the Newick tree $((U,((V,\beta,W)C),X)B)\alpha$ \\
$u,v,x,y$ & Special symbols for $U,V,X,Y$ when they are terminal strings \\
\end{tabular}
